---
title: "timeline-visualization-testing"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Purpose

To create attractive visualizations of timelines of coded behaviors from PLAY project Datavyu spreadsheets.

## Requirements

- Show mom and infant time lines by behavioral domain/code

## Generate sample data

```{r generate-data}
onset <- c(30, 45, 75, 80)
offset <- c(32, 46, 78, 90)
person <- c('infant', 'infant', 'mother', 'mother')
# Grouping by bout.id allows line segments
bout <- c(1, 2, 3, 4)

sample.1 <- data.frame(bout, person, onset, offset)
```

## Sample plot

```{r sample-plot}
sample.1 %>%
  gather(., key = event, value = time, onset, offset) %>%
  ggplot() +
  aes(x = time, y = person, group = bout) +
  geom_line(aes(size = 2)) +
  theme_classic() +
  theme(legend.position = "none")
```

Ok, this looks promising. Let's add some additional variables, including some for different 'domains' of codes.

## Sample data set 2

```{r generate-dataset-2}
onset.i <- sort(runif(n = 40, min = 1, max = 300))
onset.m <- sort(runif(n = 40, min = 1, max = 300))
offset.i <- onset.i + runif(n = 40, min = 1, max = 20)
offset.m <- onset.m + runif(n = 40, min = 1, max = 20)
onset <- c(onset.i, onset.m)
offset <- c(offset.i, offset.m)
person <- rep(c("infant", "mother"), each = 40)
bout <- 1:length(onset)
domain <- rep(c("lang", "loco", "obj", "emo"), 20)

sample.2 <- data.frame(person, domain, bout, onset, offset)
```

Now, let's plot.

```{r plot-sample-2}
sample.2 %>%
  gather(., key = event, value = time, onset, offset) %>%
  ggplot() +
  aes(x = time, y = domain, group = bout, fill = domain) +
  geom_line(aes(size = 2, color = domain)) +
  facet_grid(person ~ .) +
  theme_classic() +
  theme(legend.position = "bottom")
  
```

So, data in the following format can be easily plotted as a time line:

```{r print-data-table}
sample.2 %>%
  knitr::kable(.)
```

