---
title: "timeline-visualization-testing"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Purpose

To create attractive visualizations of timelines of coded behaviors from PLAY project Datavyu spreadsheets.

## Requirements

- Show mom and infant time lines by behavioral domain/code

## Example 1

### Generate sample data

```{r generate-data}
onset <- c(30, 45, 75, 80)
offset <- c(32, 46, 78, 90)
person <- c('infant', 'infant', 'mother', 'mother')
# Grouping by bout.id allows line segments
bout <- c(1, 2, 3, 4)

sample.1 <- data.frame(bout, person, onset, offset)
```

### Sample plot

```{r sample-plot}
sample.1 %>%
  gather(., key = event, value = time, onset, offset) %>%
  ggplot() +
  aes(x = time, y = person, group = bout) +
  geom_line(aes(size = 2)) +
  theme_classic() +
  theme(legend.position = "none")
```

Ok, this looks promising. Let's add some additional variables, including some for different 'domains' of codes.

## Example 2

### Generate data

```{r generate-dataset-2}
onset.i <- sort(runif(n = 40, min = 1, max = 300))
onset.m <- sort(runif(n = 40, min = 1, max = 300))
offset.i <- onset.i + runif(n = 40, min = 1, max = 20)
offset.m <- onset.m + runif(n = 40, min = 1, max = 20)
onset <- c(onset.i, onset.m)
offset <- c(offset.i, offset.m)
person <- rep(c("infant", "mother"), each = 40)
bout <- 1:length(onset)
domain <- rep(c("lang", "loco", "obj", "emo"), 20)

sample.2 <- data.frame(person, domain, bout, onset, offset)
```

Now, let's plot.

```{r plot-sample-2}
sample.2 %>%
  gather(., key = event, value = time, onset, offset) %>%
  ggplot() +
  aes(x = time, y = domain, group = bout, fill = domain) +
  geom_line(aes(size = 2, color = domain)) +
  facet_grid(person ~ .) +
  theme_classic() +
  theme(legend.position = "bottom")
  
```

So, data in the following format can be easily plotted as a time line:

```{r print-data-table}
sample.2 %>%
  knitr::kable(.)
```

## Example 3: Working with an actual exported Datavyu file

There are sample Datavyu files from the 'gold standard' videos (minus the emotion coding) in `analysis/csv/`. Let's see if we can use a comparable workflow to visualize these data.

```{r import-datavyu-csv}
sub.id <- 'nyu_18'
fn <- paste0("analysis/csv/", sub.id, ".csv")
dv <- read.csv(fn, stringsAsFactors = FALSE)

# Select subset of key variables, especially ordinals, onsets, & offsets
dv %>%
  select(babyloc.ordinal, babyloc.onset, babyloc.offset,
         babyobject.ordinal, babyobject.onset, babyobject.offset,
         babyvoc.ordinal, babyvoc.onset, babyvoc.offset,
         momloc.ordinal, momloc.onset, momloc.offset,
         momobject.ordinal, momobject.onset, momobject.offset,
         momspeech.ordinal, momspeech.onset, momspeech.offset) ->
  dv

# Remove NAs and make ordinals unique within domains for later grouping
dv %>%
  filter(!is.na(babyloc.ordinal), !is.na(momloc.ordinal),
         !is.na(babyobject.ordinal), !is.na(momobject.ordinal),
         !is.na(babyvoc.ordinal), !is.na(momspeech.ordinal)) %>%
  filter(!is.na(babyloc.onset), !is.na(momloc.onset),
         !is.na(babyobject.onset), !is.na(momobject.onset),
         !is.na(babyvoc.onset), !is.na(momspeech.onset)) %>%
   filter(!is.na(babyloc.offset), !is.na(momloc.offset),
         !is.na(babyobject.offset), !is.na(momobject.offset),
         !is.na(babyvoc.offset), !is.na(momspeech.offset)) ->
  dv

dv %>%
  mutate(., b.loc.bout = paste0('b.loc.', babyloc.ordinal)) %>%
  mutate(., m.loc.bout = paste0('m.loc.', momloc.ordinal)) %>%
  mutate(., b.obj.bout = paste0('b.obj.', babyobject.ordinal)) %>%
  mutate(., m.obj.bout = paste0('m.obj.', momobject.ordinal)) %>%
  mutate(., b.voc.bout = paste0('b.voc.', babyvoc.ordinal)) %>%
  mutate(., m.voc.bout = paste0('m.voc.', momspeech.ordinal)) ->
  dv1

# Gather for onsets and offsets
dv1 %>%
  gather(key = bout.lbl, value = bout, b.loc.bout, m.loc.bout, b.obj.bout, m.obj.bout,
         b.voc.bout, m.voc.bout) %>%
  arrange(bout) ->
  dv2
  
dv2 %>%
  gather(key = event, value = time.ms, 
         babyloc.onset, babyloc.offset, babyobject.onset, babyobject.offset,
         babyvoc.onset, babyvoc.offset, momloc.onset, momloc.offset, 
         momobject.onset, momobject.offset, momspeech.onset, momspeech.offset) %>%
  mutate(event = str_extract(event, "[onsetf]+$")) %>%
  mutate(person = substr(bout.lbl, 1, 1)) %>%
  mutate(domain = substr(bout.lbl, 3, 5)) %>%
  select(person, domain, bout, event, time.ms) ->
  dv2
```

Okay, now that we've munged the data, let's plot.

```{r plot-nyu-18}
dv2 %>%
  filter(time.ms < 1000*60*3) %>%
  ggplot() +
  aes(x = time.ms, y = domain, group = bout, fill = domain) +
  geom_line(aes(size = 2, color = domain)) +
  facet_grid(person ~ .) +
  theme_classic() +
  theme(legend.position = "bottom")
```

Looks like all the NA's got grouped. 

The 'elegant' way isn't working, so let's try an approach that separates the different domains.

```{r import-datavyu-segregate-by-domain}
sub.id <- 'nyu_18'
fn <- paste0("analysis/csv/", sub.id, ".csv")
dv <- read.csv(fn, stringsAsFactors = FALSE)

dv %>%
  select(., babyloc.ordinal, babyloc.onset, babyloc.offset) %>%
  gather(., key = event, value = time.ms, -babyloc.ordinal) %>%
  mutate(., event = str_extract(event, "[onsetf]+$")) %>%
  mutate(., ordinal = paste0('b.loc.', babyloc.ordinal)) %>%
  mutate(., person = "infant") %>%
  select(-babyloc.ordinal) %>%
  arrange(., time.ms) ->
  dv.babyloc

dv %>%
  select(., momloc.ordinal, momloc.onset, momloc.offset) %>%
  gather(., key = event, value = time.ms, -momloc.ordinal) %>%
  mutate(., event = str_extract(event, "[onsetf]+$")) %>%
  mutate(., ordinal = paste0('m.loc.', momloc.ordinal)) %>%
  mutate(., person = "mother") %>%
  select(-momloc.ordinal) %>%
  arrange(., time.ms) ->
  dv.momloc

dv.loc <- rbind(dv.babyloc, dv.momloc)
dv.loc$domain = "locomotion"

dv.loc %>%
  filter(time.ms < 1000*60*3) %>%
  ggplot() +
  aes(x = time.ms, y = person, group = ordinal) +
  geom_line(aes(size = 2))
```

```{r dfs-for-other-domains}
dv %>%
  select(., babyobject.ordinal, babyobject.onset, babyobject.offset) %>%
  gather(., key = event, value = time.ms, -babyobject.ordinal) %>%
  mutate(., event = str_extract(event, "[onsetf]+$")) %>%
  mutate(., ordinal = paste0('b.obj.', babyobject.ordinal)) %>%
  mutate(., person = "infant") %>%
  select(-babyobject.ordinal) %>%
  arrange(., time.ms) ->
  dv.babyobj

dv %>%
  select(., momobject.ordinal, momobject.onset, momobject.offset) %>%
  gather(., key = event, value = time.ms, -momobject.ordinal) %>%
  mutate(., event = str_extract(event, "[onsetf]+$")) %>%
  mutate(., ordinal = paste0('m.obj.', momobject.ordinal)) %>%
  mutate(., person = "mother") %>%
  select(-momobject.ordinal) %>%
  arrange(., time.ms) ->
  dv.momobj

dv.obj <- rbind(dv.babyobj, dv.momobj)
dv.obj$domain = "objects"

dv %>%
  select(., babyvoc.ordinal, babyvoc.onset, babyvoc.offset) %>%
  gather(., key = event, value = time.ms, -babyvoc.ordinal) %>%
  mutate(., event = str_extract(event, "[onsetf]+$")) %>%
  mutate(., ordinal = paste0('b.voc.', babyvoc.ordinal)) %>%
  mutate(., person = "infant") %>%
  select(-babyvoc.ordinal) %>%
  arrange(., time.ms) ->
  dv.babyvoc

dv %>%
  select(., momspeech.ordinal, momspeech.onset, momspeech.offset) %>%
  gather(., key = event, value = time.ms, -momspeech.ordinal) %>%
  mutate(., event = str_extract(event, "[onsetf]+$")) %>%
  mutate(., ordinal = paste0('m.voc.', momspeech.ordinal)) %>%
  mutate(., person = "mother") %>%
  select(-momspeech.ordinal) %>%
  arrange(., time.ms) ->
  dv.momspeech

dv.voc <- rbind(dv.babyvoc, dv.momspeech)
dv.voc$domain = "speech"
```

Let's plot the separate domains.

```{r plot-obj-speech}
dv.obj %>%
  filter(time.ms < 1000*60*5) %>%
  ggplot() +
  aes(x = time.ms, y = person, group = ordinal) +
  geom_line(aes(size = 2))

dv.voc %>%
  filter(time.ms < 1000*60*3) %>%
  ggplot() +
  aes(x = time.ms, y = person, group = ordinal) +
  geom_point() +
  geom_line(aes(size = 2))
```

Infant and mother vocalization and speech events have offset == onset!

Putting it all together.

```{r}
dv.nyu18 <- rbind(dv.loc, dv.obj, dv.voc)
dv.nyu18 %>%
  arrange(time.ms) ->
  dv.nyu18

dv.nyu18 %>%
  filter(time.ms > 1000*60*30, time.ms < 1000*60*32) %>%
  ggplot() +
  aes(x = (time.ms)/(60*1000), y = person, group = ordinal, color = domain) +
  geom_point() +
  geom_line(aes(size = 1)) +
  facet_grid(domain ~ .) +
  xlab("Minutes") +
  theme_classic()
```

Ok, so this is a start. I'd still like to find a more `tidyverse` compatible way to do the munging. But sometimes a hack is the best way to start.
