---
title: "Video segment extraction"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  html_document:
      code_folding: show
      toc: true
      toc_float: true
params:
  db_login_id: "email@provider.com"
---

# Purpose

This notebook documents Rick Gilmore's explorations of ways to extract video segments from coded PLAY Project Datavyu files.

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages
library(tidyverse)
```

# Goals

1. Select a coding pass for a PLAY participant.
1. Open the Datavyu file.
1. Extract the segments for a particular code value.
1. Convert the segments into URLs.
1. Select some of the segments and return the videos from Databrary.

# Technical strategy

## Available `.opf` files

Login to Databrary.

```{r}
databraryapi::login_db(params$db_login_id)
```

The PLAY pilot directory is volume 444. 
List the sessions in this volume

```{r}
play_vol <- 444
(play_pilot_sessions <- databraryapi::list_sessions(play_vol))
```

Filter out Materials folders.
The `top` variable is TRUE for Materials folders.

```{r}
play_pilot_sessions <- play_pilot_sessions %>%
  dplyr::filter(., is.na(top))
play_pilot_sessions
```

There are `r dim(play_pilot_sessions)[1]` pilot sessions.
I had planned to look at each individually, but I am not sure that any of these have `.opf` files.
So, for now I'm going to ask the staff where the files might be.

In the meantime, I have the file for NYU-018 locally.
I have uploaded it to Databrary.

I need to identify the session ID for participant 18.

```{r}
play_pilot_018 <- play_pilot_sessions %>%
  dplyr::filter(., stringr::str_detect(name, "018"))
play_pilot_018
```

Then list the assets in the session.

```{r}
play_pilot_018_assets <- databraryapi::list_assets_in_session(play_pilot_018$session_id)
```

Note that the Datavyu file has mimetype `application/vnd.datavyu` and extention `opf`.

## Download, clean `opf`

Create the download directory if needed.

```{r}
dv_dir = 'opf'
if (!dir.exists(dv_dir)) {
  dir.create(dv_dir)
}
```

Then download the file.

```{r}
play_018_dv <- play_pilot_018_assets %>%
  dplyr::filter(., mimetype == 'application/vnd.datavyu')
```


```{r, eval=FALSE}
databraryapi::download_datavyu(session_id = play_pilot_018$session_id, asset_id = play_018_dv$asset_id, out_dir = dv_dir)
```

Now, let's convert the Datavyu file to a CSV.

```{r, eval=FALSE}
databraryapi::extract_dv(in_dir = dv_dir)
databraryapi::dv_to_csv(dv_dir)
```

That seems to work.

## Examine codes

Let's examine the various code types available.
First, we import the CSV as a data frame.

```{r}
dv_fn <- list.files(path = dv_dir, pattern = "\\.csv$", full.names = TRUE)
# Need to read times as character 'c' to then convert fractions of a second properly.
p_018 <- readr::read_csv(dv_fn, col_types = 'cccc')  
unique(p_018$code)
```

## Focus on `momspeech`

Let's focus on the `momspeech` code for now.

```{r}
momtalk <- p_018 %>%
  dplyr::filter(., code == 'momspeech')
momtalk
```

Let's convert these onsets and offsets into URIs that point to the video on Databrary.

The format of a video URI is as follows:

`https://nyu.databrary.org/slot/<SLOT_ID>/<SEG_START_MS>,<SEG_END_MS>/asset/<ASSET_ID>/download?inline=true`

We'll have to decide the default duration for the speech events since these are point cell events.
Also, I wonder if we should have a cushion at the start.

### Determine video

```{r}
play_pilot_018_assets
```

Because the videos have an idioysyncractic naming convention, I have to choose the correct one manually.
Here it is `S#18_1-Hour.mov`.

```{r}
p_018_video_asset <- play_pilot_018_assets %>%
  dplyr::filter(., name == 'S#18_1-Hour.mov')
video_asset_id <- p_018_video_asset$asset_id
```


```{r}
# Databrary 
segment_range <- databraryapi::get_asset_segment_range(vol_id = play_vol, session_id = play_pilot_018$session_id, asset_id = play_018_dv$asset_id)

# helper function
HHMMSSmmm_to_ms <- function(HHMMSSmmm = "01:01:01.333") 
{
    if (!is.character(HHMMSSmmm)) {
        stop("HHMMSSmmm must be a string.")
    }
    if (stringr::str_detect(HHMMSSmmm, "([0-9]{2}):([0-9]{2}):([0-9]{2})\\.([0-9]{3})")) {
        time_segs <- stringr::str_match(HHMMSSmmm, "([0-9]{2}):([0-9]{2}):([0-9]{2})\\.([0-9]{3})")
        as.numeric(time_segs[5]) + as.numeric(time_segs[4]) * 
            1000 + as.numeric(time_segs[3]) * 1000 * 60 + as.numeric(time_segs[2]) * 
            1000 * 60 * 60
    }
    else {
        NULL
    }
}

code_to_URI <- function(code_row, slot_id, asset_id, pre_offset_ms = -500, post_offset_ms = 1500) {
  # code_row is a code from a Datavyu data.frame/tibble
  # TODO: Add parameter checks
  
  # Process onsets & offsets
  onset_ms <- HHMMSSmmm_to_ms(as.character(code_row$onset)) + pre_offset_ms
  offset_ms <- HHMMSSmmm_to_ms(as.character(code_row$offset)) + post_offset_ms
  selected_seg <- paste0(onset_ms, ",", offset_ms)
  
  # Create URI
  paste0('https://nyu.databrary.org/slot/', slot_id, '/', selected_seg, 
         '/asset/', asset_id, '/download?inline=true')
}
```

Shall we try this with a few segments?

```{r}
code_to_URI(momtalk[10,], slot_id = play_pilot_018$session_id, asset_id = video_asset_id)
code_to_URI(momtalk[24,], slot_id = play_pilot_018$session_id, asset_id = video_asset_id)
code_to_URI(momtalk[34,], slot_id = play_pilot_018$session_id, asset_id = video_asset_id)
code_to_URI(momtalk[740,], slot_id = play_pilot_018$session_id, asset_id = video_asset_id)
```

That seems to work.
Let's modify the `code_to_URI()` function to make it `lapply()` or `purr()` compatible.

```{r}
code_to_URI_2 <- function(i, df, slot_id, asset_id, pre_offset_ms = -500, post_offset_ms = 1500) {
  # code_row is a code from a Datavyu data.frame/tibble
  # TODO: Add parameter checks
  code_row <- df[i,]
  
  # Process onsets & offsets
  onset_ms <- HHMMSSmmm_to_ms(as.character(code_row$onset)) + pre_offset_ms
  offset_ms <- HHMMSSmmm_to_ms(as.character(code_row$offset)) + post_offset_ms
  selected_seg <- paste0(onset_ms, ",", offset_ms)
  
  # Create URI
  paste0('https://nyu.databrary.org/slot/', slot_id, '/', selected_seg, 
         '/asset/', asset_id, '/download?inline=true')
}
```

```{r}
uri_list <- lapply(1:dim(momtalk)[1], code_to_URI_2, df = momtalk, slot_id = play_pilot_018$session_id, asset_id = video_asset_id)
```

### PLAY_018 mom utterances with links

```{r}
momtalk_w_uris <- momtalk %>%
  dplyr::mutate(., uri = unlist(uri_list))

momtalk_w_uris %>%
  dplyr::select(., code.value, uri) %>%
  knitr::kable(.)
```

Ok, so this works.

### Logout

```{r}
databraryapi::logout_db()
```

## Shiny app

The following is merely a template app.
I have set `eval=FALSE` for now.

```{r, echo=FALSE, eval=FALSE}
shiny_dir <- "shiny/"
shiny_app <- "databrary"
shinyAppDir(appDir = paste(shiny_dir, shiny_app, sep=""),
  options=list(
    width="100%", height=700
  )
)
```

